---
title: "HW2_SDS315"
author: "Anh Doan(atd2354)"
output:   
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(stringr)

profs <- read.csv("profs.csv")
bike <- read.csv("bikeshare.csv")
capmetro_UT <- read.csv("capmetro_UT.csv")
billboard <- read.csv("billboard.csv")

profs_ggplot <- ggplot(profs)
bike_ggplot <- ggplot(bike)
capmetro_UT_ggplot <- ggplot(capmetro_UT)
billboard_ggplot <- ggplot(billboard)
```

# **Problem 1: Beauty, or not, in the classroom** 

### **Part A: Distribution of Course Evaluation Scores**

```{r, echo = FALSE}
profs_ggplot + geom_histogram(aes(x=eval), fill = "darkred", col = "black") + labs(title = "Distribution of Course Evaluation Scores", x = "Course Evaluation Scores", y = "Number of Professors") + ylim(0,40) 
```

### **Part B: Distribution of Course Evaluation Scores by Native English Proficiency**

```{r}
profs_ggplot + geom_boxplot(aes(x=native, y = eval, fill= native)) +
  theme(legend.position = "none") +
  labs(title = "Distribution of Evaluation Scores by Native English Proficiency",
       x = "Is the Professor a Native English Speaker?",
       y = "Course Evaluation Scores")
```


### **Part C: Distribution of Course Evaluation Scores by Gender**

```{r}
profs_ggplot + geom_histogram(aes(x=eval, fill = gender), col = "black") +
  facet_wrap(~gender, nrow=2) +
  theme(strip.text = element_text(size = 12, face = "bold"),
        legend.position = "none") + 
  labs(title = "Distribution of Course Evaluation Scores by Gender", 
       x = "Course Evaluation Scores",
       y = "Number of Professors") 
```

### **Part D: Physical Attractiveness vs Evaluation Scores**

```{r}
profs_ggplot + geom_point(aes(x=eval, y=beauty), alpha = 0.6, color = "red") +
  geom_smooth(method = "lm", aes(x=eval, y = beauty), se = FALSE) +
  labs(title = "Relationship Between Physical Attractiveness and Evaluation Scores",
       x = "Course Evaluation Scores",
       y = "Physical Attractiveness")
```

# **Problem 2: Bike Sharing**

### **Part A: Average Hourly Bike Rentals**

```{r}
avg1 <- aggregate(total ~ hr, bike, mean)

ggplot(avg1) + geom_line(aes(x=hr, y=total)) +
  geom_point(aes(x=hr, y=total),color = "red", size = 2) +
  labs(
    title = "Average Hourly Bike Rentals",
    x = "Hour of the Day",
    y = "Average Bike Rentals"
  ) +
  scale_x_continuous(breaks = 0:23) 
```

### **Part B: Average Hourly Bike Rentals on Workdays vs Weekends**

```{r}
avg2 <- aggregate(total ~ hr + workingday, bike, mean)

ggplot(avg2) + geom_line(aes(x=hr, y = total, group = workingday)) +
  facet_wrap(~workingday, labeller = as_labeller(c("0" = "Non-Working Day", "1" = "Working Day"))) +
  geom_point(aes(x=hr, y=total),color = "red", size = 2) +
  theme(strip.text = element_text(face = "bold", size = 14)) +
  labs(title = "Hourly Bike Rentals: Working vs Non-Working Days",
      x = "Hour of the Day",
      y = "Average Bike Rentals") 
```

### **Part C: 9 AM Bike Rentals by Weather and Workday Status**

```{r}
subset <- bike[bike$hr == 9,]
avg3 <- aggregate(total~weathersit + workingday, subset, mean)

ggplot(avg3, aes(x = weathersit, y = total, fill = weathersit)) +
  geom_bar(stat = "identity", color = "black") +
  labs(
    title = "Average 9 AM Bike Rentals by Weather Situation and Workday Status",
    x = "Weather Situation",
    y = "Average Bike Rentals"
  ) +
  facet_wrap(~ workingday, labeller = as_labeller(c("0" = "Non-Working Day", "1" = "Working Day"))) +
  theme(strip.text = element_text(face = "bold", size = 13), legend.position = "none") 
```

# **Problem 3: Capital Metro UT Ridership**

```{r}
capmetro_UT <- mutate(capmetro_UT, day_of_week = factor(day_of_week,levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")), month = factor(month,levels=c("Sep", "Oct","Nov")))

avg4 <- aggregate(boarding ~ hour_of_day + day_of_week + month, capmetro_UT, mean)
avg_ggplot <- ggplot(avg4)
```

### **Part 1: Average Boardings by Hour of The Day, Day of The Week, and Month**

```{r}
avg_ggplot + geom_line(aes(x=hour_of_day, y = boarding, color = month)) +
  facet_wrap(~day_of_week,) +
  labs(
    title = "Average Boardings by Hour of The Day, Day of The Week, and Month",
    x = "Hours of The Day",
    y = "Average Boardings",
    color = "Month"
  ) +
  theme(
    strip.text = element_text(size = 12, face = "bold")) +
  scale_x_continuous(
    breaks = seq(0, 23, by = 2)
  )
```

### **Part 2: Boardings vs Temperature Faceted by Hour of The Day**

```{r}
#Convert variables to upper case
capmetro_UT$weekend <- str_to_title(capmetro_UT$weekend)
capmetro_UT_ggplot <- ggplot(capmetro_UT)

capmetro_UT_ggplot + geom_point(aes(x=temperature, y=boarding, color = weekend)) +
  facet_wrap(~hour_of_day) +
  labs(
    title = "Boardings vs Temperature Faceted by Hour of The Day",
    x = "Temperature",
    y = "Boardings",
    color = "Day Type"
  )
```

# **Problem 4: Wrangling the Billboard Top 100**

### **Part A: Top 10 Most Popular Songs Since 1958**

```{r}
grouped <- group_by(billboard, song)
filtered <- filter(grouped, weeks_on_chart == max(weeks_on_chart))

# Summarize the total weeks for each unique song
summarized <- summarise(filtered, Performer = first(performer), count = max(weeks_on_chart))

# Sort in descending order of count
sorted <- summarized[order(-summarized$count),]
head(sorted, 10)
```

### **Part B: Musical Diversity in the Billboard Top 100 Over Time**

```{r, results="hide"}
filtered <- filter(billboard, year!= 1958 & year!= 2021)
diversity <- summarise(group_by(filtered, year), count = length(unique(song)))

diversity_ggplot <- ggplot(diversity)
diversity_ggplot + 
  geom_point(aes(x=year, y=count), color="red") +
  geom_line(aes(x=year, y =count)) +
  labs(
    title = "Musical Diversity in the Billboard Top 100 Over Time",
    x = "Year",
    y = "Number of Songs"
  ) +
   scale_x_continuous(
    breaks = seq(1960,2020, by = 5))
```

### **Part C: Artists With More Than 30 Ten-Week Hits**

```{r}
# Filter for songs that appeared on the Billboard Top 100 for at least 10 weeks
ten_week_hits <- billboard %>%
  group_by(performer, song) %>%
  summarize(weeks_on_chart = n(), .groups = "drop") %>%
  filter(weeks_on_chart >= 10) %>%
  group_by(performer) %>%
  summarize(ten_week_hits = n(), .groups = "drop") %>%
  filter(ten_week_hits >= 30)

# Create the bar plot
ggplot(ten_week_hits, aes(x = reorder(performer, ten_week_hits), y = ten_week_hits)) +
  geom_bar(stat = "identity", fill = "darkred", color = "black") +
  coord_flip() +
  labs(title = "Artists with More Than 30 Ten-Week Hits",
       x = "Artist",
       y = "Number of Ten-Week Hits") 
```

